---
title: "Empirical Analysis of Airbnb Listing Prices in Boston, USA"
author: "Riley Fewer, Polyna German, Arya Badal Kalarickal, Omar Salaheldin Shahata"
date: "April 3, 2026"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
library(readr)
library(ggplot2)
library(patchwork)
library(knitr)
library(car)

# Load and prepare data
df <- as.data.frame(
  read_csv("data/cleaned/listings_cleaned.csv") %>%
    mutate(
      property_type = factor(property_type,
                             levels = c("Apartment", "Condominium",
                                        "House", "Other")),
      host_is_superhost = factor(
        ifelse(host_is_superhost == TRUE, "t", "f"),
        levels = c("f", "t"))
    )
)
```
## Section 1: Introduction

Short-term rental platforms such as Airbnb have significantly impacted 
tourism and housing markets. In cities like Boston, where housing supply 
is limited but demand is high due to tourism and overall desirability, 
understanding what drives variation in Airbnb listing prices is both 
relevant and important for hosts, guests, and policymakers. Airbnb 
pricing influences host revenue, market accessibility, and the 
relationship between short-term rentals and long-term housing 
availability. As a result, identifying the factors associated with 
Airbnb prices contributes to broader discussions in real estate and 
urban economics.

This project investigates the following research question: **To what 
extent do property type, capacity, and listing characteristics explain 
variation in the log-transformed nightly price of Airbnb listings in 
Boston?** The response variable of interest is the logarithm of the 
nightly listing price. Log transformation is used because nightly 
Airbnb prices exhibit a strongly right-skewed distribution and taking 
the logarithm of price helps better satisfy the linear regression 
assumptions of linearity, constant variance, and approximately normally 
distributed residuals. The preliminary model included six predictors: 
number of bedrooms, number of bathrooms, number of guests accommodated, 
property type, security deposit, and superhost status. These variables 
capture key dimensions of physical size, guest capacity, and listing 
characteristics. Through model selection, superhost status was 
subsequently removed, yielding a final model with five predictors.

Existing academic literature provides motivation for this analysis. 
Prior research demonstrates a strong relationship between property size, 
capacity, and prices. Cai et al. (2019) finds that nightly Airbnb 
prices in Hong Kong increase with property size and capacity, reporting 
that entire homes and private rooms are priced 377% higher than shared 
rooms and that each additional bedroom increases price by approximately 
17.8%. Similarly, Gibbs et al. (2018) finds that private rooms are 
priced 28.3% and 40.1% lower than entire homes in Toronto and Calgary, 
respectively, highlighting the premium hosts place on privacy. 

Linear regression is well-suited for this analysis because it allows 
for the simultaneous examination of multiple numerical and categorical 
predictors while maintaining interpretability. Furthermore, the 
continuous nature of the log-transformed response variable supports 
meaningful interpretation of estimated coefficients as percentage 
changes in price.

## Section 2: Data Description

The dataset used for our research project - Boston Airbnb Open Data 
published on Kaggle (Boston Airbnb Open Data, n.d.) originates from 
Inside Airbnb, a project aimed at providing data on Airbnb's impact on 
residential communities (Inside Airbnb, n.d). For this project, we used the publicly available data provided to determine how property 
characteristics explain variation in the log-transformed nightly price 
of Airbnb listings in Boston. The data was originally gathered by 
web-scraping Airbnb listings pages (Inside Airbnb, n.d.).

The response variable is the logarithm of the nightly listing price 
(log\_price). The log transformation is used because prices exhibit a 
strongly right-skewed distribution (Figure 1A), causing both 
non-normality and heteroskedasticity. Therefore, taking the logarithm 
of price better satisfies the linear regression assumptions of linearity, 
constant variance, and approximately normally distributed residuals, 
making log\_price a suitable continuous response variable for our linear 
regression model (Figure 1B). Of the 3,585 raw listings, 2,243 had 
missing values for security_deposit, which were set to zero on the basis 
that a missing deposit indicates no requirement, consistent with Airbnb listing practice. After removing observations with missing values for 
price, bedrooms, or bathrooms, the cleaned dataset contains 3,561 
observations.

The preliminary model included six predictor variables: number of 
bedrooms, number of bathrooms, guests accommodated, security deposit, 
property type, and superhost status. Of these, superhost status was 
removed during model selection in Section 4, resulting in a final model with five predictors.

```{r fig1, fig.cap="A) Raw nightly price distribution showing strong right skew. B) Log-transformed price distribution showing approximate symmetry, justifying the log transformation.", fig.width=9, fig.height=4}
p1a <- ggplot(df, aes(x = price)) +
  geom_histogram(bins = 60, fill = "steelblue", colour = "white") +
  labs(title = "A. Price Distribution (Raw)",
       x = "Nightly Price ($)", y = "Frequency") +
  xlim(0, 1500) +
  theme_bw()

p1b <- ggplot(df, aes(x = log_price)) +
  geom_histogram(bins = 40, fill = "steelblue", colour = "white") +
  labs(title = "B. Log Price Distribution",
       x = "log(Nightly Price)", y = "Frequency") +
  theme_bw()

p1a + p1b
```

Figure 2A shows that the majority of listings do not charge a security 
deposit. Figure 2B shows a highly right-skewed distribution of non-zero 
deposits, with the majority of amounts being between \$100 and \$500. 
According to Ert et al. (2016), listings with higher value and quality 
often include deposits, which indirectly correlate with higher pricing.

```{r fig2, fig.cap="A) Count of listings with and without a security deposit. The majority require no deposit. B) Distribution of non-zero deposit amounts, showing right skew with most values between $100 and $500.", fig.width=9, fig.height=4}
p2a <- ggplot(df,
              aes(x = factor(ifelse(security_deposit == 0,
                                    "No Deposit ($0)",
                                    "Deposit (>$0)")))) +
  geom_bar(fill = c("steelblue", "firebrick"), alpha = 0.8) +
  labs(title = "A. Deposits non-zero vs zero",
       x = "", y = "Count of listings") +
  theme_bw()

p2b <- ggplot(subset(df, security_deposit > 0),
              aes(x = security_deposit)) +
  geom_histogram(fill = "steelblue", colour = "white", bins = 30) +
  labs(title = "B. Distribution of non-zero deposits",
       x = "Amount ($)", y = "Count") +
  theme_bw()

p2a + p2b
```

Figures 3A, 3B, and 3D show the distributions of bedrooms, bathrooms, 
and guests\_included. These variables are discrete and right-skewed. 
According to Wang and Nicolau (2017), larger properties with more 
bedrooms and bathrooms affect nightly rates due to increased comfort and 
accommodation capacity. Figure 3C shows that superhost status is rare. 
Superhosts can charge relatively high prices because the badge functions 
as a quality signal that reduces uncertainty and increases consumer trust 
(Airbnb, n.d.; Almeida et al., 2025). Figure 3E displays the major trend of property 
type to be an apartment. Amenities and increased privacy play a major 
role in pricing according to Gibbs et al. (2018).

```{r fig3, fig.cap="Distributions of all predictor variables. A) Bedrooms, B) Bathrooms, C) Superhost status, D) Guests included, E) Property type. All continuous predictors are right-skewed. Apartments dominate the property type distribution.", fig.width=11, fig.height=7}
p3a <- ggplot(df, aes(x = factor(bedrooms))) +
  geom_bar(fill = "steelblue") +
  labs(title = "A. Bedrooms", x = "Count", y = "Frequency") +
  theme_bw()

p3b <- ggplot(df, aes(x = factor(bathrooms))) +
  geom_bar(fill = "steelblue") +
  labs(title = "B. Bathrooms", x = "Count", y = "Frequency") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p3c <- ggplot(df, aes(x = host_is_superhost)) +
  geom_bar(fill = "steelblue") +
  scale_x_discrete(labels = c("f" = "No", "t" = "Yes")) +
  labs(title = "C. Superhost Status",
       x = "Is Superhost", y = "Frequency") +
  theme_bw()

p3d <- ggplot(df, aes(x = factor(guests_included))) +
  geom_bar(fill = "steelblue") +
  labs(title = "D. Guests Included", x = "Count", y = "Frequency") +
  theme_bw()

p3e <- ggplot(df, aes(y = property_type)) +
  geom_bar(fill = "steelblue") +
  labs(title = "E. Property Type", x = "Frequency", y = NULL) +
  theme_bw()

(p3a + p3b + p3c) / (p3d + p3e)
```

## Section 3: Primary Model Results and Diagnostics
```{r primary-model}
# Fit primary model
model_primary <- lm(
  log_price ~ bedrooms + bathrooms + guests_included +
    property_type + host_is_superhost + security_deposit,
  data = df
)

summary_primary <- summary(model_primary)
```

The preliminary linear regression estimated the logarithmic nightly 
prices of Airbnb listings in Boston using the six predictors mentioned 
above. The initial model explains a meaningful portion of price variation 
with an adjusted $R^2$ of 0.273. The joint significance of the predictors 
is confirmed by the F-statistic of 167.77 ($p < 0.001$).

Consistent with hedonic pricing models (Gibbs et al., 2018; Wang & 
Nicolau, 2017), larger listings have higher rates. Each additional 
bedroom and bathroom is associated with a 32.9% and 14.8% increase in nightly price respectively. The bedroom effects align with Cai's (2019) findings of a 17.8% increase in Hong Kong, though our stronger estimate 
likely reflects Boston's premium urban landscape. Guest capacity also has 
a significantly positive effect, consistent with literature. Security 
deposits exhibit a small but statistically significant positive 
association with price, serving as a signal for listing quality (Ert et 
al., 2016). Among the categorical variables, property type plays a 
significant role. Houses are priced lower than apartments, contrasting 
with Gibbs et al.'s (2018) findings in Toronto and Calgary where entire 
homes commanded premiums. This reversal could be attributed to Boston's 
apartment-centric urban core. Superhost status is not statistically 
significant in the preliminary model, differing from Almeida et al. 
(2025) and suggesting physical characteristics dominate reputational 
signals in Boston's mature market.

```{r table1, echo=FALSE}
# Coefficient table
coef_df <- as.data.frame(summary_primary$coefficients)
ci       <- confint(model_primary)

coef_table <- data.frame(
  Predictor = c("Intercept", "Bedrooms", "Bathrooms",
                "Guests Included", "Condominium vs Apt",
                "House vs Apt", "Other vs Apt",
                "Superhost (t vs f)", "Security Deposit"),
  Estimate   = round(coef_df[, "Estimate"], 4),
  Std_Error  = round(coef_df[, "Std. Error"], 4),
  t_value    = round(coef_df[, "t value"], 3),
  p_value    = round(coef_df[, "Pr(>|t|)"], 4)
)
rownames(coef_table) <- NULL

kable(coef_table,
      col.names = c("Predictor", "Estimate", "Std. Error",
                    "t-value", "p-value"),
      caption = "Primary model regression coefficients.
      Reference category for property type is Apartment.",
      digits = 4)
```

```{r table2, echo=FALSE}
# Model fit statistics
fit_table <- data.frame(
  Statistic = c("R-squared", "Adjusted R-squared",
                "F-statistic", "p-value", "N"),
  Value = c(
    round(summary_primary$r.squared, 4),
    round(summary_primary$adj.r.squared, 4),
    round(summary_primary$fstatistic[1], 2),
    "< 0.001",
    nrow(df)
  )
)

kable(fit_table,
      col.names = c("Statistic", "Value"),
      caption = "Primary model fit statistics.")
```

Regression diagnostics indicate linear model assumptions are reasonable 
but imperfect. The residuals are clustered around zero, suggesting that 
linearity holds well, though slight curvature exhibits mild 
heteroskedasticity for higher fitted values. The Scale-Location plot 
confirms some variance instability. The Q-Q plot reveals upper-tail 
deviations, indicating mild departures from normality. Given the large 
sample size, these minor violations are unlikely to substantially bias 
estimates, but motivated the model refinements described in Section 4.

```{r fig4, fig.cap="Residual diagnostic plots for the primary model. Left: Residuals vs Fitted showing approximate linearity with mild heteroskedasticity at higher fitted values. Centre: Q-Q plot showing mild upper-tail deviation from normality. Right: Scale-Location plot confirming mild variance instability.", fig.width=10, fig.height=3.5}
par(mfrow = c(1, 3))
plot(model_primary, which = 1,
     main = "Residuals vs Fitted", sub = "")
plot(model_primary, which = 2,
     main = "Q-Q Residuals", sub = "")
plot(model_primary, which = 3,
     main = "Scale-Location", sub = "")
par(mfrow = c(1, 1))
```

## Section 4: Model Selection
```{r section4-setup, include=FALSE}
# Fit transformation comparison model (on full data, for transformation decision)
model_log_deposit <- lm(
  log_price ~ bedrooms + bathrooms + guests_included +
    property_type + host_is_superhost + log1p(security_deposit),
  data = df
)

# Log transform of bedrooms and bathrooms (for transformation decision)
model_log_rooms <- lm(
  log_price ~ log1p(bedrooms) + log1p(bathrooms) + guests_included +
    property_type + host_is_superhost + security_deposit,
  data = df
)

# Power transform confirmation (on full data)
pt_result <- car::powerTransform(
  price ~ bedrooms + bathrooms + guests_included +
    property_type + host_is_superhost + security_deposit,
  data = df
)

# Cook's Distance and influential observation removal (from primary model)
cooks_d         <- cooks.distance(model_primary)
threshold       <- 4 / nrow(df)
influential_idx <- which(cooks_d > threshold)
df_final        <- df[-influential_idx, ]

# Refit full model on cleaned data (post-outlier removal)
model_no_inf <- lm(
  log_price ~ bedrooms + bathrooms + guests_included +
    property_type + host_is_superhost + security_deposit,
  data = df_final
)

# Stepwise selection on cleaned data (CORRECT order: outliers removed first)
model_step_aic <- step(model_no_inf, direction = "backward",
                       trace = 0)
model_step_bic <- step(model_no_inf, direction = "backward",
                       trace = 0, k = log(nrow(df_final)))

# AIC of model_no_inf vs model_no_inf without superhost (for prose reporting)
model_no_superhost_clean <- lm(
  log_price ~ bedrooms + bathrooms + guests_included +
    property_type + security_deposit,
  data = df_final
)

aic_with    <- AIC(model_no_inf)
aic_without <- AIC(model_no_superhost_clean)
bic_with    <- BIC(model_no_inf)
bic_without <- BIC(model_no_superhost_clean)

# VIF on the final predictor set (no superhost), clean data
vif_vals <- car::vif(model_no_superhost_clean)
```

The final model was reached by three main decisions: Box-Cox 
transformation, outlier handling, and variable selection.

The log transformation of the response variable (nightly price) was 
formally confirmed using a Box-Cox power transformation analysis, which 
returned $\lambda = 0.039$. Since $\lambda \approx 0$ corresponds to a 
log transformation, this provides statistical justification for the 
choice motivated in Section 2 by the right-skewed raw price distribution 
(Figure 1). For the continuous predictors, two transformation candidates 
were evaluated using AIC as the criterion. First, replacing 
security\_deposit with $\log(1 + \text{security\_deposit})$, which 
handles the large number of zero-deposit listings (as can be seen in 
Figure 2A), was tested against the primary model. The transformed model 
produced an AIC of 5977.72, compared to 5942.12 for the untransformed 
version, an increase of 35.6 points. The Scale-Location plots for both 
specifications were visually identical (Figure 5), confirming no 
heteroskedasticity improvement. Second, applying $\log(1 + x)$ to both 
bedrooms and bathrooms was tested; this worsened the AIC by 131.0 points 
(AIC = 6073.12 versus 5942.12). In conclusion, no predictor 
transformations were applied, so all of the predictors used in the model 
are on their original scale. The decision to not transform predictors 
also keeps the straightforward percentage-change interpretation of 
coefficients in the log-response model.

```{r fig5, fig.cap="Scale-Location plots comparing the primary model using raw security deposit (left) and log1p(security deposit) (right). The smoother trend and residual spread are nearly identical across both plots, providing no evidence that the transformation improves homoscedasticity.", fig.width=9, fig.height=4}
par(mfrow = c(1, 2))
plot(model_primary,     which = 3,
     main = "Scale-Location: Raw Deposit", sub = "")
plot(model_log_deposit, which = 3,
     main = "Scale-Location: log1p(Deposit)", sub = "")
par(mfrow = c(1, 1))
```

The primary model's Cook's Distance plot (Figure 6) and influence plot 
(Figure 7) were used to identify observations that disproportionately 
affected the fitted coefficients. Leverage values were computed and 
compared against the Hoaglin and Welsch (1978) threshold of 
$2(p+1)/n = 2(9)/3{,}561 \approx 0.005$, where $p = 8$ counts all 
predictors excluding the intercept, and $n = 3{,}561$ is the full 
cleaned sample. Cook's Distance was then computed for every observation, 
using the standard threshold of $4/n = 4/3{,}561 \approx 0.00112$. In 
total, 161 observations exceeded this threshold and were removed. Among 
the most influential were observations 2614 (Cook's $D = 0.0139$), 212 
(Cook's $D = 0.0100$), 3032 (Cook's $D = 0.0041$), and 2264 (Cook's 
$D = 0.0027$). To be precise, observations 2264 and 3032 correspond to 
one- and two-bedroom Boston apartments listed at \$4,000 and \$3,000 per 
night respectively, prices far from market normal that generated 
standardized residuals exceeding $+4.9$ and $+4.5$, making these points 
clear outliers. Observations 373 and 1416, which appear in the influence 
plot (Figure 7) as the rightmost points with hat-values of 0.090 and 
0.070, were examined but not removed. Since their Cook's $D$ values 
(0.00043 and 0.00019) fall below the removal threshold, and their 
standardized residuals are small enough ($-0.20$ and $-0.15$), 
indicating that even though they show extreme values, they do not distort 
the actual prediction. After removing the 161 influential observations, 
the cleaned dataset for the final model contains $n = 3{,}400$ 
observations. The Cook's Distance plot for the final model (Figure 9) 
confirms the success of these removals: the maximum Cook's $D$ falls 
from 0.0139 to 0.0022, and the final residual diagnostics (Figure 8) 
show improved symmetry in the Residuals vs. Fitted plot and better Q-Q 
alignment relative to Figure 4.

```{r fig6, fig.cap="Cook's Distance for all observations in the primary model. The dashed red line marks the threshold of 4/n = 0.00112. Observations 212, 1062, and 2614 are the most influential.", fig.width=8, fig.height=4.5}
plot(model_primary, which = 4,
     main = "Cook's Distance (Primary Model)", sub = "")
abline(h = threshold, col = "firebrick", lty = 2)
```

```{r fig7, echo=FALSE, results='hide', fig.cap="Influence plot showing hat-values (leverage) on the x-axis, studentized residuals on the y-axis, and Cook's Distance encoded as circle size and colour. Observations 373 and 1416 are high leverage but not influential. Observations 2264 and 3032 are extreme outliers in Y. Observations 212 and 2614 are genuinely influential.", fig.width=8, fig.height=7}
car::influencePlot(model_primary)
```

In order to select variables we used backward stepwise selection, 
utilizing AIC and BIC as different scoring methods. Results of both 
indicated that dropping the host\_is\_superhost predictor would improve 
the model. Removing this variable resulted in a reduction of AIC by 
`r round(aic_with - aic_without, 2)` points and BIC by 
`r round(bic_with - bic_without, 2)`, meaning that this predictor just 
added complexity to the model without improving fit. None of the other 
predictors were removed: bedrooms, bathrooms, guests included, property 
type, and security deposit were all beneficial to the model, meaningfully 
contributing to fit. To confirm the absence of multicollinearity, VIF 
were computed for all explanatory variables; all of them were below 1.52, 
indicating that our coefficient estimates are not biased.

## Section 5: Final Model Inference and Results
```{r section5-setup, include=FALSE}
# Fit final model on cleaned data
model_final <- lm(
  log_price ~ bedrooms + bathrooms + guests_included +
    property_type + security_deposit,
  data = df_final
)

summary_final <- summary(model_final)
```

The final model regresses log-transformed nightly prices on five 
predictors: bedroom count, bathroom count, guests included, property 
type, and security deposit. The model was fit on 3,400 observations 
after removal of influential points. Table 3 presents the estimated 
coefficients, standard errors, 95% confidence intervals, and p-values.

```{r table5, echo=FALSE}
coef_df <- as.data.frame(summary_final$coefficients)
ci      <- confint(model_final, level = 0.95)

final_table <- data.frame(
  Predictor  = c("Intercept", "Bedrooms", "Bathrooms",
                 "Guests Included", "Condominium vs Apt",
                 "House vs Apt", "Other vs Apt",
                 "Security Deposit"),
  Estimate   = round(coef_df[, "Estimate"], 4),
  Std_Error  = round(coef_df[, "Std. Error"], 4),
  CI_Lower   = round(ci[, 1], 4),
  CI_Upper   = round(ci[, 2], 4),
  t_value    = round(coef_df[, "t value"], 3),
  p_value    = round(coef_df[, "Pr(>|t|)"], 4),
  Pct_Change = round((exp(coef_df[, "Estimate"]) - 1) * 100, 2)
)
rownames(final_table) <- NULL

kable(final_table,
      col.names = c("Predictor", "Estimate", "Std. Error",
                    "95% CI Lower", "95% CI Upper",
                    "t-value", "p-value", "% Change"),
      caption   = "Final model coefficients, standard errors,
      95% confidence intervals, p-values, and percentage change in
      nightly price. Reference category for property type is Apartment.",
      digits    = 4)
```

Note: The intercept percentage change (7629.71%) represents the 
predicted nightly price when all predictors equal zero and has no 
meaningful real-world interpretation.

**Coefficient Interpretations.** For a log-response model, a coefficient 
$\hat{\beta}$ represents the estimated average change in $\log(\text{price})$ 
for a one-unit increase in the predictor, holding all other predictors 
constant. The corresponding percentage change in nightly price is 
$(\exp(\hat{\beta}) - 1) \times 100\%$.

Each additional bedroom is associated with a 30.6% increase in nightly 
price (95% CI: 26.8% to 34.5%, $p < 0.001$), holding all other 
predictors constant. Bedroom count is associated with the largest 
positive effect among the continuous predictors and directly addresses 
the capacity dimension of the research question. Each additional bathroom 
is independently associated with a 17.2% price increase (95% CI: 12.0% 
to 22.6%, $p < 0.001$). The independence of this effect suggests that 
guests value bathroom availability as a distinct comfort signal beyond 
sleeping capacity. Each additional increase in guest capacity is 
associated with an 8.9% increase in price (95% CI: 7.0% to 10.9%, 
$p < 0.001$), capturing a third capacity signal that operates 
independently of physical room count. Bedroom count and guest capacity 
capture different aspects of a listing: guest capacity represents a hard 
upper limit on how many people can be accommodated, whereas bedroom count 
reflects how that capacity is distributed across the listing. 
Collectively, the three capacity predictors address the research 
question's capacity dimension: physical size and accommodation capacity 
explain a meaningful and consistent portion of price variation in Boston 
Airbnb listings.

With respect to property type, houses are priced 41.0% lower than 
comparable apartments (95% CI: $-43.8\%$ to $-38.0\%$, $p < 0.001$). 
This is the single largest effect in the final model. Listings in the 
Other category are priced 11.2% below apartments ($p = 0.011$). 
Condominiums show no statistically significant price difference from 
apartments ($\hat{\beta} = 0.006$, $p = 0.872$, 95% CI spanning zero), 
indicating Boston's condominium market is priced on par with apartments 
when physical characteristics are held constant. This non-result is 
itself informative: the structural similarity between apartments and 
condominiums in an urban market means property type distinction between 
these two carries no independent pricing signal. In contrast, the 
significantly lower pricing of houses suggests a clear urban preference 
for apartments and condominiums in Boston's Airbnb market.

Focusing on listing characteristics, each additional dollar of security 
deposit is associated with a 0.03% increase in nightly price 
($p < 0.001$). While statistically significant, the economic magnitude 
is negligible, a \$100 deposit increase corresponds to only a 2.9% 
price increase. Consistent with Ert et al. (2016), this likely reflects 
security deposit as a proxy for listing quality rather than a direct 
pricing mechanism: hosts who charge deposits tend to manage 
higher-quality listings, and the deposit captures that latent quality 
signal indirectly.

**Literature Comparison.** The bedroom coefficient of 30.6% per bedroom 
substantially exceeds the 17.8% reported by Cai et al. (2019) for Hong 
Kong listings. This divergence is interpretable rather than surprising: 
Boston's premium housing market, where residential space is scarce and 
expensive, means that each additional private bedroom commands a 
disproportionately higher return than in higher-density Asian cities 
where space constraints are distributed differently across property 
types. The finding aligns directionally with Cai et al. (2019) that
bedrooms matter, while the magnitude reflects Boston's specific market 
conditions.

The house discount of 41.0% relative to apartments is directionally 
consistent with Gibbs et al. (2018), who found private rooms and houses 
priced 28.3% to 40.1% below entire homes in Toronto and Calgary. The 
estimate from the final model falls at the upper end of this range, 
consistent with the apartment-dominated composition of Boston's 
short-term rental market, where 73.4% of listings are apartments. Our 
finding reinforces Gibbs et al.'s (2018) core conclusion that property 
type is a primary determinant of Airbnb pricing across North American 
urban markets, with the specific magnitude reflecting the relative 
scarcity of apartments versus houses in Boston's listing inventory.

The non-significance of superhost status ($p = 0.548$, dropped in 
Section 4) diverges from Almeida et al. (2025), who found review-based 
quality signals to be significant price determinants. This divergence is 
consistent with the interpretation offered in the proposal: in Boston's 
established Airbnb market, guests rely on physical property attributes 
rather than host reputation cues when assessing value. The security 
deposit finding aligns with Ert et al. (2016) in showing that 
listing-level quality signals operate through pricing via indirect 
proxies rather than host-level reputation mechanisms.

**Model Performance.** Table 4 summarises performance metrics for both 
the primary and final models.

```{r table6, echo=FALSE}
perf_table <- data.frame(
  Metric = c("R-squared", "Adjusted R-squared", "AIC", "BIC",
             "Residual Std. Error", "F-statistic", "p-value", "N"),
  Primary_Model = c(
    round(summary_primary$r.squared, 4),
    round(summary_primary$adj.r.squared, 4),
    round(AIC(model_primary), 2),
    round(BIC(model_primary), 2),
    round(summary_primary$sigma, 4),
    round(summary_primary$fstatistic[1], 2),
    "< 2.2e-16",
    nrow(df)
  ),
  Final_Model = c(
    round(summary_final$r.squared, 4),
    round(summary_final$adj.r.squared, 4),
    round(AIC(model_final), 2),
    round(BIC(model_final), 2),
    round(summary_final$sigma, 4),
    round(summary_final$fstatistic[1], 2),
    "< 2.2e-16",
    nrow(df_final)
  )
)

kable(perf_table,
      col.names = c("Metric", "Primary Model", "Final Model"),
      caption   = "Model performance comparison between the
      primary model (6 predictors, n = 3,561) and the final model
      (5 predictors, n = 3,400).")
```

The final model achieves an adjusted $R^2$ of 0.3013, explaining 30.1% 
of the variation in log nightly price. The improvement over the primary 
model's adjusted $R^2$ of 0.2726 is attributable primarily to the removal 
of influential observations rather than the change in predictors, which 
was the expected result, as influential observations inflate residual 
variance and suppress $R^2$. Residual standard error decreased from 
0.5565 to 0.5118, indicating a tighter and more accurate fit around the 
regression line. The F-statistic of 210.37 ($p < 2.2 \times 10^{-16}$) 
confirms that the model as a whole is highly significant, the five 
predictors collectively explain a statistically significant portion of 
log-price variation. The AIC of 5103.43 and BIC of 5158.61 are lower 
than the primary model values, though the direct numerical comparison is 
approximate since the models were fit on different sample sizes following 
outlier removal. An adjusted $R^2$ of 0.30 is consistent with the range 
reported in the hedonic pricing literature for Airbnb datasets, Gibbs 
et al. (2018) and Wang and Nicolau (2017) report comparable values, and 
reflects the reality that structural variables alone cannot capture the 
full complexity of short-term rental pricing, where unobserved factors 
such as neighbourhood, interior quality, entertainment amenities, and 
listing photographs contribute substantially to price variation. The 
final model's diagnostic plots (Figure 8) show no systematic pattern in 
residuals versus fitted values, supporting the linearity assumption, and 
the Q-Q plot shows only mild upper-tail deviation consistent with a small 
number of high-priced listings that the model moderately underpredicts.

```{r fig8-final, fig.cap="Residual diagnostic plots for the final model. Left: Residuals vs Fitted showing no systematic pattern, supporting linearity. Centre: Q-Q plot showing mild upper-tail deviation consistent with a small number of high-priced listings. Right: Scale-Location showing improved homoscedasticity relative to the primary model.", fig.width=10, fig.height=3.5}
par(mfrow = c(1, 3))
plot(model_final, which = 1,
     main = "Residuals vs Fitted (Final)", sub = "")
plot(model_final, which = 2,
     main = "Q-Q Residuals (Final)", sub = "")
plot(model_final, which = 3,
     main = "Scale-Location (Final)", sub = "")
par(mfrow = c(1, 1))
```

```{r fig9-final-cooks, fig.cap="Cook's Distance for the final model. The maximum Cook's D has fallen from 0.0139 in the primary model to 0.0022, confirming the success of the influential observation removal.", fig.width=8, fig.height=4.5}
plot(model_final, which = 4,
     main = "Cook's Distance (Final Model)", sub = "")
abline(h = 4 / nrow(df_final), col = "firebrick", lty = 2)
```



\pagebreak

## Section 6: Discussion and Conclusion

This study examined the extent to which property type, capacity, and 
listing characteristics explain variation in the log-transformed nightly 
price of Airbnb listings in Boston. After a rigorous model selection 
process, encompassing Box-Cox transformation verification, Cook's 
Distance based outlier removal, and backward stepwise variable selection, 
the final linear regression model was fit on 3,400 observations with 
five predictors: bedroom count, bathroom count, guest capacity, property 
type, and security deposit. The model achieves an adjusted $R^2$ of 
0.3013 and an F-statistic of 210.37 ($p < 2.2 \times 10^{-16}$), 
confirming that the predictors collectively explain a statistically 
significant and meaningful portion of log-price variation. The central 
conclusion of this analysis is that physical capacity is the dominant 
driver of Airbnb nightly prices in Boston, consistent with hedonic 
pricing theory and prior literature.

The most substantive finding concerns the magnitude and independence of 
three capacity predictors. Each additional bedroom is associated with a 
30.6% increase in nightly price, ceteris paribus (95% CI: 26.8% to 
34.5%, $p < 0.001$). This estimate exceeds the 17.8% per-bedroom premium 
reported by Cai et al. (2019) in Hong Kong, a difference that can be 
explained by Boston's premium urban real estate market and the relative 
scarcity of larger multi-bedroom listings. Similarly, each additional 
bathroom is associated with a 17.2% price increase (95% CI: 12.0% to 
22.6%, $p < 0.001$), reflecting the value guests assign to comfort and 
privacy. Guest accommodation capacity also exhibits a significantly 
positive association with price, consistent with Wang and Nicolau's 
(2017) analysis showing that larger properties command higher rates. Each 
unit increase in guest capacity is associated with an 8.9% price increase 
(95% CI: 7.0% to 10.9%, $p < 0.001$). Taken together, these 
capacity-related findings confirm that physical scale is a primary driver 
of Airbnb pricing in Boston, in line with the broader hedonic pricing 
literature (Gibbs et al., 2018).

The role of property type reveals an interesting pattern in the Boston 
market. Houses are priced 41.0% lower than apartments (95% CI: $-43.8\%$ to $-38.0\%$, $p < 0.001$),
and listings in the Other category are priced 
11.2% below apartments ($p = 0.011$). The house discount aligns with 
Gibbs et al.'s (2018) finding that houses and private rooms are priced 
28.3% to 40.1% below entire homes in Toronto and Calgary. Our 41.0% 
estimate sits at the upper end of this range, consistent with Boston's 
apartment-dominated nature, where 73.4% of listings are apartments. This 
finding reflects the broader conclusion that property type is a primary 
determinant of Airbnb pricing across North American urban markets (Gibbs 
et al., 2018), with the specific magnitude reflecting Boston's relative 
scarcity of desirable non-apartment stock in high-demand tourist areas.

The non-significance of superhost status is another notable finding. 
Although Almeida et al. (2025) demonstrates that review-based quality 
signals, including superhost designation, positively influence Airbnb 
prices in some markets, this effect does not emerge in our Boston dataset. 
Superhost status was removed from the model through backward stepwise 
selection, as dropping it reduced AIC by 1.59 points and BIC by 7.72, 
showing that it added complexity without improving model fit. One 
interpretation is that Boston's short-term rental market is sufficiently 
mature that guests weigh physical amenities more heavily than reputational 
badges when making decisions. Alternatively, superhost status may have 
indirect pricing effects that are mediated by listing quality indicators 
already captured by other predictors in the model. Additionally, security 
deposit retains a small but statistically significant positive association 
with price, though its economic magnitude is negligible, a \$100 increase 
in deposit corresponds to only a 2.9% price increase. As Ert et al. 
(2016) argued, deposits serve as a proxy signal for listing quality 
rather than a direct pricing mechanism, as hosts managing higher-quality 
listings tend to charge deposits.

The final model's adjusted $R^2$ of 0.3013 represents a meaningful 
improvement over the primary model's 0.2726, primarily due to the removal 
of 161 influential observations that were inflating residual variance. 
The residual standard error fell from 0.5565 to 0.5118, indicating a 
tighter fit around the regression line. AIC and BIC are both lower than 
in the primary model. Our final adjusted $R^2$ of 30.1% is consistent 
with comparable literature in hedonic pricing, while remaining variation 
is attributable to unobserved factors such as neighbourhood, 
desirability, seasonal demand, and interior quality. The diagnostic plots 
confirm that linearity holds and the Q-Q plot shows only mild upper-tail 
deviation, consistent with a small number of high-priced listings that 
the model underpredicts.

Based on these findings, several recommendations emerge for hosts, guests, 
and policymakers. For hosts, the results indicate that investing in 
additional bedrooms and increasing physical capacity has the strongest 
potential to justify higher nightly prices. Hosts seeking to maximise 
revenue should prioritise listing features that increase physical capacity, 
as these have the largest reliable association with price. For guests, 
these findings highlight that pricing in Boston reflects tangible amenities 
more than reputational signals, suggesting that careful attention to 
physical listing details may yield better value than relying on superhost 
status alone. From a policy perspective, the strong relationship between 
property capacity and price is relevant to discussions about Airbnb's 
effect on housing markets. As Barron et al. (2020) demonstrate, higher 
Airbnb activity is associated with elevated housing costs; the fact that 
larger units command higher premiums may incentivise landlords to convert 
family-sized units into short-term rentals.

There are several ways in which this analysis could be strengthened in 
future studies. First, the current model explains only 30.1% of price 
variation, indicating that important predictors are absent. Incorporating 
neighbourhood-level variables such as proximity to transit, tourist 
attractions, or high-demand areas could improve explanatory power. 
Spatial regression techniques or neighbourhood fixed effects would allow 
the model to account for geographic clustering of high-priced listings. 
Second, the dataset is cross-sectional, which prevents analysis of how 
pricing varies by season, major events, or changes in the broader rental 
market. A longitudinal design incorporating time fixed effects would 
enable more nuanced causal inference. Third, the inclusion of review 
scores, review count, and host response rate as additional predictors 
would allow a more direct test of whether reputation-based signals 
influence pricing in Boston, rather than relying on superhost status as 
a blunt proxy. Finally, expanding the sample to include listings from 
multiple cities, following the approach of Wang and Nicolau (2017), 
would allow for cross-market comparisons that could illuminate whether 
Boston's distinctive patterns, the lower relative pricing of houses and 
the insignificance of superhost status, generalise to other urban 
markets.

In conclusion, this study provides empirical evidence that physical 
capacity characteristics, specifically bedroom count, bathroom count, 
and guest accommodation, are the primary drivers of nightly price 
variation among Boston Airbnb listings. These results are broadly 
consistent with hedonic pricing literature while also revealing 
market-specific patterns, most notably the strong house discount relative 
to apartments and the limited pricing influence of superhost status. 
These findings contribute to a growing body of evidence that Airbnb 
pricing reflects underlying property attributes in predictable ways, 
while demonstrating that city-level context meaningfully shapes which 
attributes matter most. Future work incorporating reputation-based 
variables and neighbourhood-level controls has strong potential to 
deepen this understanding and narrow the gap between model predictions 
and observed listing prices.

## Section 7: Author Contributions

**Omar Salaheldin Shahata:** Co-authored Section 3 (Primary Model Results and 
Diagnostics), including statistical interpretation of the preliminary 
model coefficients and residual diagnostic analysis. Co-authored Section 
5 (Final Model Inference and Results), contributing to the coefficient 
table presentation and model performance discussion. Contributed in the final 
formatting of the report, including ensuring word counts are met and 
references cited correctly. 

**Arya Badal Kalarickal:** Developed and maintained all R scripts for 
the final analysis pipeline, including data cleaning, exploratory data 
analysis, primary model fitting, diagnostic analysis, model selection, 
and final model inference. Contributed to the writing of Section 3 
(Primary Model) and Section 4 (Model Selection), including the 
influential observation analysis and variable selection procedures. Wrote 
Section 5 (Final Model Inference and Results), including all coefficient 
interpretations, literature comparisons, and model performance assessment.

**Polyna German:** Authored the introduction of the dataset and predictor 
variable descriptions, and created all corresponding visualizations in 
Section 2 (Data Description). Contributed to the writing of Section 4 
(Model Selection), including the influential observation analysis, outlier 
identification and justification using Cook's Distance, interpretation of 
variable selection results, referencing and interpreting created graphs, 
and revising the statistical writing for precision and consistency. Also 
contributed to the References section.

**Riley Fewer:** Authored the Introduction (Section 1), including the 
research question, literature review connecting three peer-reviewed 
sources to the research question, and justification for linear regression 
as the analytical approach. Also authored the Discussion and Conclusion 
(Section 6), synthesizing key findings, situating results within the 
existing literature, providing recommendations for hosts, guests, and 
policymakers, and outlining directions for future research.

\pagebreak

## Section 8: References

Almeida, A., Nunes, A. P., & Machado, L. P. (2025). How do reviews 
impact Airbnb's prices? A hedonic approach. *Tourism and Hospitality*, 
*6*(4), 181. https://doi.org/10.3390/tourhosp6040181

Barron, K., Kung, E., & Proserpio, D. (2020). The effect of home-sharing 
on house prices and rents: Evidence from Airbnb. *Marketing Science*, 
*40*(1), 23--47. https://doi.org/10.1287/mksc.2020.1227

Cai, Y., Zhou, Y., Ma, J., & Scott, N. (2019). Price determinants of 
Airbnb listings: Evidence from Hong Kong. *Tourism Analysis*, *24*(2), 
227--242. https://doi.org/10.3727/108354219x15525055915554

Ert, E., Fleischer, A., & Magen, N. (2016). Trust and reputation in the 
sharing economy: The role of personal photos in Airbnb. *Tourism 
Management*, *55*, 62--73. 
https://doi.org/10.1016/j.tourman.2016.01.013

Gibbs, C., Guttentag, D., Gretzel, U., Morton, J., & Goodwill, A. 
(2018). Pricing in the sharing economy: A hedonic pricing model applied 
to Airbnb listings. *Journal of Travel & Tourism Marketing*, *35*(1), 
46--56. https://doi.org/10.1080/10548408.2017.1308292

Wang, D., & Nicolau, J. L. (2017). Price determinants of sharing economy 
based accommodation rental: A study of listings from 33 cities on 
Airbnb.com. *International Journal of Hospitality Management*, *62*, 
120--131. https://doi.org/10.1016/j.ijhm.2016.12.007

Boston Airbnb Open Data. (n.d.). *Boston Airbnb open data* [Data set]. 
Kaggle. Retrieved February 1, 2026, from 
https://www.kaggle.com/datasets/airbnb/boston

Inside Airbnb. (n.d.). *About Inside Airbnb: Mission and data collection 
description*. Retrieved February 1, 2026, from 
https://insideairbnb.com/about/

Inside Airbnb. (n.d.). *Data assumptions: How the data were compiled and 
processed*. Retrieved February 1, 2026, from 
https://insideairbnb.com/data-assumptions/

Airbnb. (n.d.). *What's required to be a Superhost*. Retrieved February 
1, 2026, from https://www.airbnb.ca/help/article/829
